---
title: "SFDS Final Project"
author: "Adam E."
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Packages
```{r Load_libraries, include = FALSE, echo = FALSE} 
# install.packages("multcomp")
#install.packages("xtable")
#install.packages("vcd")

# Required libraries
library(tidyverse)
library(broom)
library(gridExtra)
#library(pwr)
#library(emmeans)
#library(GGally)
library(car)

# library(maps)
# library(kableExtra)
# library(class)
# library(patchwork)
# library(RCurl) 
# library(jsonlite)

#Import all libraries - ggplot2 for visualization
library(ggplot2)
library(ggthemes)
library(klaR)
library(caret)
# library(e1071)
# library(vcd)


```


# Data Import from csv
```{r csv_import}
# Read in Data
Path <- "https://raw.githubusercontent.com/Weiprecht/Weiprecht.github.io/main/HousingData/train.csv"
AmesHousing_Data = read.csv(Path, header=TRUE)

#MF = read.csv(file.choose(), header = TRUE)

# Drop rows where 'Id' column contains "MM"
AmesHousing_Data <- AmesHousing_Data[!grepl( "NA", AmesHousing_Data$Id), ]


AmesHousing_Data2 <- AmesHousing_Data %>%
  filter(Neighborhood %in% c("NAmes", "Edwards", "BrkSide"))
AmesHousing_Data2$Neighborhood <- as.factor(AmesHousing_Data2$Neighborhood)

# Create a new column that holds the'SalePrice' in hundreds
AmesHousing_Data2$SalePriceHundr <- floor(AmesHousing_Data2$SalePrice /100)

# Create a new column that holds the'SalePrice' in hundreds
AmesHousing_Data2$GrLivAreaHundr <- floor(AmesHousing_Data2$GrLivArea /100)

# Write new dataset
# write.csv(AmesHousing_Data2, "D:/University/SMU/Statistical_Foundations_for_Data_Science/Stat_foundations_repository/Statistical_Foundations/Final_Project/AmesHousing_Data2.csv", row.names = FALSE)

AmesHousing_Data2

```
# Initial dataset for Analysis Question 1: check assumptions
```{r housing_regression}

# Fit a multiple linear regression model
lm_model <- lm(SalePriceHundr ~ GrLivAreaHundr + Neighborhood, data = AmesHousing_Data2)

# Summarize the model
summary(lm_model)

# Calculate VIF for the variables in the model
vif_result <- vif(lm_model)
vif_result

# Calculate mean and standard deviation
mean_sale_price <- mean(AmesHousing_Data2$SalePriceHundr)
sd_sale_price <- sd(AmesHousing_Data2$SalePriceHundr)
mean_sqr_ft <- mean(AmesHousing_Data2$GrLivAreaHundr)
sd_sqr_ft <- sd(AmesHousing_Data2$GrLivAreaHundr)

# Check Assumptions Visually
plot(lm_model)
avPlots(lm(SalePriceHundr ~ GrLivAreaHundr + Neighborhood, data = AmesHousing_Data2))  

# For histograms
ggplot(AmesHousing_Data2, aes(x = SalePriceHundr)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "grey") +
  geom_density(aes(color = "Actual")) +
  stat_function(aes(color = "Normal"), 
                fun = dnorm, 
                args = list(mean = mean_sale_price, sd = sd_sale_price)) +
  xlab("Sale Price (in Hundreds)") +
  scale_colour_manual("Distribution", values = c("darkred", "darkblue"))

ggplot(AmesHousing_Data2, aes(x = GrLivAreaHundr)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "grey") +
  geom_density(aes(color = "Actual")) +
  stat_function(aes(color = "Normal"), 
                fun = dnorm, 
                args = list(mean = mean_sqr_ft, sd = sd_sqr_ft)) +
  xlab("Area sqr ft (in Hundreds)") +
  scale_colour_manual("Distribution", values = c("darkred", "darkblue"))

# Boxplot for SalePrice 
ggplot(AmesHousing_Data2, aes(x = Neighborhood, y = SalePriceHundr)) +
  geom_boxplot() +
  labs(title = "SalePrice Distribution Across Neighborhoods",
       x = "Neighborhood",
       y = "SalePrice (in hundreds)")

# Boxplot for square feet 
ggplot(AmesHousing_Data2, aes(x = Neighborhood, y = GrLivAreaHundr)) +
  geom_boxplot() +
  labs(title = "Square Feet Across Neighborhoods",
       x = "Neighborhood",
       y = "Square Feet (in tens)")

```
# Since the is some right skewedness and variables that are leveraging the3 regression - look at the outliers
```{r examine_outliers}
# Calculate IQR for SalePrice and GrLivArea
Q1 <- quantile(AmesHousing_Data2$SalePrice, 0.25)
Q3 <- quantile(AmesHousing_Data2$SalePrice, 0.75)
IQR_SalePrice <- Q3 - Q1

Q1_area <- quantile(AmesHousing_Data2$GrLivArea, 0.25)
Q3_area <- quantile(AmesHousing_Data2$GrLivArea, 0.75)
IQR_GrLivArea <- Q3_area - Q1_area

# Define thresholds for outliers 
threshold_SalePrice <- 2.5 * IQR_SalePrice
threshold_GrLivArea <- 2.5 * IQR_GrLivArea

# Identify outliers
outliers <- AmesHousing_Data2[(AmesHousing_Data2$SalePrice < (Q1 - threshold_SalePrice)) |
                              (AmesHousing_Data2$SalePrice > (Q3 + threshold_SalePrice)) |
                              (AmesHousing_Data2$GrLivArea < (Q1_area - threshold_GrLivArea)) |
                              (AmesHousing_Data2$GrLivArea > (Q3_area + threshold_GrLivArea)), ]
outliers

# Assuming 'AmesHousing_Data2' is your dataset
partial_sales <- AmesHousing_Data2$SaleCondition == "Partial"
partial_rows <- AmesHousing_Data2[partial_sales, ]

partial_rows

# Export outliers to a CSV file
# write.csv(outliers, "D:/University/SMU/Statistical_Foundations_for_Data_Science/Stat_foundations_repository/Statistical_Foundations/Final_Project/outliers.csv", row.names = FALSE)
```



# 524(131) and 1299(339) in Edwards, have over 10000 sqft including the unfinished basements and 1 and 1.5 acres-most likley an error - no homes over .5 acres built after 2007
# 534(136) in Brookside, has a price and square feet that are likely missing a zero - Brookside smallest structure has 729 square feet this one, 334
# 725(190) in Edwards, 643 (169) in NAmes prices are high for living space and thus the sales-living regression is off
```{r outliers_high_leverage}
# Identify the rows of interest'
outlier_indices <- c( 131, 136, 169, 190, 339)

# Extracting rows corresponding to the outlier indices
outlier_rows <- AmesHousing_Data2[outlier_indices, ]

outlier_rows


```
```{r drop_investigated_outliers}
# Drop rows 131, 136, and 339
rows_to_drop <- c( 131, 136, 190, 339)
AmesHousing_Data3 <- AmesHousing_Data2[-rows_to_drop, ]

AmesHousing_Data3
summary(AmesHousing_Data3)

# Fit a multiple linear regression model
lm_reduced <- lm(SalePriceHundr ~ GrLivAreaHundr + Neighborhood, data = AmesHousing_Data3)

# Summarize the model
summary(lm_reduced)

# Calculate VIF for the variables in the model
vif_reduced <- vif(lm_reduced)
vif_reduced


# Calculate mean and standard deviation
mean_sale_price <- mean(AmesHousing_Data3$SalePriceHundr)
sd_sale_price <- sd(AmesHousing_Data3$SalePriceHundr)
mean_sqr_ft <- mean(AmesHousing_Data3$GrLivAreaHundr)
sd_sqr_ft <- sd(AmesHousing_Data3$GrLivAreaHundr)

# Check Assumptions Visually
plot(lm_reduced)
avPlots(lm(SalePriceHundr ~ GrLivAreaHundr + Neighborhood, data = AmesHousing_Data3))

# Identify outliers
identify(lm_reduced$fit, labels = row.names(AmesHousing_Data3), plot = TRUE)

# For histograms
ggplot(AmesHousing_Data3, aes(x = SalePriceHundr)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "grey") +
  geom_density(aes(color = "Actual")) +
  stat_function(aes(color = "Normal"), 
                fun = dnorm, 
                args = list(mean = mean_sale_price, sd = sd_sale_price)) +
  xlab("Sale Price (in Hundreds)") +
  scale_colour_manual("Distribution", values = c("darkred", "darkblue"))

ggplot(AmesHousing_Data3, aes(x = GrLivAreaHundr)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "grey") +
  geom_density(aes(color = "Actual")) +
  stat_function(aes(color = "Normal"), 
                fun = dnorm, 
                args = list(mean = mean_sqr_ft, sd = sd_sqr_ft)) +
  xlab("Area sqr ft (in Hundreds)") +
  scale_colour_manual("Distribution", values = c("darkred", "darkblue"))

# Boxplot for SalePrice 
ggplot(AmesHousing_Data3, aes(x = Neighborhood, y = SalePriceHundr)) +
  geom_boxplot() +
  labs(title = "SalePrice Distribution Across Neighborhoods",
       x = "Neighborhood",
       y = "SalePrice (in hundreds)")

# Boxplot for square feet 
ggplot(AmesHousing_Data3, aes(x = Neighborhood, y = GrLivAreaHundr)) +
  geom_boxplot() +
  labs(title = "Square Feet Across Neighborhoods",
       x = "Neighborhood",
       y = "Square Feet (in tens)")

```


```{r internal_cross_validation}
# Define the model formula
formula <- SalePriceHundr ~ GrLivAreaHundr + Neighborhood

# Create a train control object for cross-validation
ctrl <- trainControl(method = "cv",  # Cross-validation method
                     number = 5,     # Number of folds (you can change this)
                     verboseIter = TRUE,  # Display iteration progress
                     summaryFunction = defaultSummary)  # Use default summary function

# Train the model with cross-validation
lm_model <- train(formula,
                  data = AmesHousing_Data2,
                  method = "lm",  # Linear regression method
                  trControl = ctrl)

# View results
lm_model
```






```{r logged_housing_data}
# Create a new column called 'lSalePrice' and place the log of the values in 'SalePrice' in this column
AmesHousing_Data3$lSalePriceHundr <- log(AmesHousing_Data3$SalePriceHundr)

# Create a new column called 'lGrLivArea' and place the log of the values in 'GrLivArea' in this column
AmesHousing_Data3$lGrLivAreaHundr <- log(AmesHousing_Data3$GrLivAreaHundr)

AmesHousing_Data3

# Write new dataset
#write.csv(AmesHousing_Data3, "D:/Website/Weiprecht.github.io/HousingData/AmesHousing_Data3.csv", row.names = FALSE)
```



```{r housing_regression_logged}

# Fit a multiple linear regression model
lm_model2 <- lm(lSalePriceHundr ~ lGrLivAreaHundr + Neighborhood, data = AmesHousing_Data3)

# Summarize the model
summary(lm_model2)

# Get coefficients and statistical summary in a tidy format
tidy_summary2 <- tidy(lm_model2)
print(tidy_summary2)

plot(lm_model2)
```

```{r internal_cross_validation2}
# Define the model formula
formula <- lSalePriceHundr ~ lGrLivAreaHundr + Neighborhood

# Create a train control object for cross-validation
ctrl <- trainControl(method = "cv",  # Cross-validation method
                     number = 5,     # Number of folds (you can change this)
                     verboseIter = TRUE,  # Display iteration progress
                     summaryFunction = defaultSummary)  # Use default summary function

# Train the model with cross-validation
lm_model <- train(formula,
                  data = AmesHousing_Data3,
                  method = "lm",  # Linear regression method
                  trControl = ctrl)

# View results
lm_model

```



```{r housing_regression_logged2}
# ref is always the element that does not show up in the model.
cat("\n-------------------------------Neighborhood no interaction----------------------\n")
pricefit = lm(lSalePriceHundr ~ relevel(Neighborhood, ref = "BrkSide") + lGrLivAreaHundr, data = AmesHousing_Data3)
summary(pricefit)

confint(pricefit)

# Get coefficients and statistical summary in a tidy format
tidy_summary2 <- tidy(pricefit)
print(tidy_summary2)

plot(pricefit)

cat("\n--------------------------Neighborhood interaction---------------------\n")

# No Interaction (Parallel Regression Lines)
pricefit2 = lm(lSalePriceHundr ~ relevel(Neighborhood, ref = "BrkSide") * lGrLivAreaHundr, data = AmesHousing_Data3)
summary(pricefit2)

confint(pricefit2)

# Get coefficients and statistical summary in a tidy format
tidy_summary2 <- tidy(pricefit2)
print(tidy_summary2)

plot(pricefit2)
```



```{r histogram_comparisons}

# Filter the data for each neighborhood
NAmes_data <- filter(AmesHousing_Data3, Neighborhood == "NAmes")
Edwards_data <- filter(AmesHousing_Data3, Neighborhood == "Edwards")
BrkSide_data <- filter(AmesHousing_Data3, Neighborhood == "BrkSide")

Combined_plot <- ggplot() +
  geom_histogram(data = NAmes_data, aes(x = lSalePriceHundr, fill = "NAmes"), color = "black", alpha = 0.5, bins = 30, show.legend = TRUE) +
  geom_histogram(data = Edwards_data, aes(x = lSalePriceHundr, fill = "Edwards"), color = "black", alpha = 0.5, bins = 30, show.legend = TRUE) +
  geom_histogram(data = BrkSide_data, aes(x = lSalePriceHundr, fill = "BrkSide"), color = "black", alpha = 0.5, bins = 30, show.legend = TRUE) +
  labs(title = "Sale Prices by Neighborhood",
       x = "Sale Price (in Hundreds)",
       y = "Frequency") +
  scale_fill_manual(values = c("NAmes" = "cornflowerblue", "Edwards" = "red", "BrkSide" = "green"), 
                    name = "Neighborhood") +
 guides(fill = guide_legend(title = "Neighborhood"))
# Show legend
#Combined_plot + guides(fill = guide_legend(title = "Neighborhood"))

#-------------------------NAmes-------------------------------------

# Calculate mean and standard deviation
mean_sale_price <- mean(NAmes_data$lSalePriceHundr)
sd_sale_price <- sd(NAmes_data$lSalePriceHundr)

# Create histogram with normal and density curves (NAMES)
NAmes_plot <- ggplot(NAmes_data, aes(x = lSalePriceHundr)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "cornflowerblue") +
  geom_density(aes(color = "Actual")) +
  stat_function(aes(color = "Normal"), 
                fun = dnorm, 
                args = list(mean = mean_sale_price, sd = sd_sale_price)) +
  ggtitle("Sale Prices in North Ames") +
  xlab("Sale Price (in Hundreds)") +
  scale_colour_manual("Distribution", values = c("darkred", "darkblue"))

#--------------------------Edwards---------------------------------------

# Calculate mean and standard deviation
mean_sale_price <- mean(Edwards_data$lSalePriceHundr)
sd_sale_price <- sd(Edwards_data$lSalePriceHundr)

# Create histogram with normal and density curves (NAMES)
Edwards_plot <- ggplot(Edwards_data, aes(x = lSalePriceHundr)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "red") +
  geom_density(aes(color = "Actual")) +
  stat_function(aes(color = "Normal"), 
                fun = dnorm, 
                args = list(mean = mean_sale_price, sd = sd_sale_price)) +
  ggtitle("Sale Prices in Edwards") +
  xlab("Sale Price (in Hundreds)") +
  scale_colour_manual("Distribution", values = c("darkred", "darkblue"))

#--------------------------BrkSide--------------------------------
# Calculate mean and standard deviation
mean_sale_price <- mean(BrkSide_data$lSalePriceHundr)
sd_sale_price <- sd(BrkSide_data$lSalePriceHundr)

# Create histogram with normal and density curves (BrkSide)
BrkSide_plot <- ggplot(BrkSide_data, aes(x = lSalePriceHundr)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "green") +
  geom_density(aes(color = "Actual")) +
  stat_function(aes(color = "Normal"), 
                fun = dnorm, 
                args = list(mean = mean_sale_price, sd = sd_sale_price)) +
  ggtitle("Sale Prices in BrkSide") +
  xlab("Sale Price (in Hundreds)") +
  scale_colour_manual("Distribution", values = c("darkred", "darkblue"))

grid.arrange(Combined_plot, NAmes_plot, Edwards_plot, BrkSide_plot, ncol = 2, nrow = 2)
```


```{r histogram_comparisons2}

# Filter the data for each neighborhood
NAmes_data <- filter(AmesHousing_Data3, Neighborhood == "NAmes")
Edwards_data <- filter(AmesHousing_Data3, Neighborhood == "Edwards")
BrkSide_data <- filter(AmesHousing_Data3, Neighborhood == "BrkSide")

Combined_plot <- ggplot() +
  geom_histogram(data = NAmes_data, aes(x = lGrLivAreaHundr, fill = "NAmes"), color = "black", alpha = 0.5, bins = 30, show.legend = TRUE) +
  geom_histogram(data = Edwards_data, aes(x = lGrLivAreaHundr, fill = "Edwards"), color = "black", alpha = 0.5, bins = 30, show.legend = TRUE) +
  geom_histogram(data = BrkSide_data, aes(x = lGrLivAreaHundr, fill = "BrkSide"), color = "black", alpha = 0.5, bins = 30, show.legend = TRUE) +
  labs(title = "Square Feet by Neighborhood",
       x = "Square Feet (in hundreds)",
       y = "Frequency") +
  scale_fill_manual(values = c("NAmes" = "cornflowerblue", "Edwards" = "red", "BrkSide" = "green"), 
                    name = "Neighborhood") +
 guides(fill = guide_legend(title = "Neighborhood"))
# Show legend
#Combined_plot + guides(fill = guide_legend(title = "Neighborhood"))

#-------------------------NAmes-------------------------------------

# Calculate mean and standard deviation
mean_Square_Feet <- mean(NAmes_data$lGrLivAreaHundr)
sd_Square_Feet <- sd(NAmes_data$lGrLivAreaHundr)

# Create histogram with normal and density curves (NAMES)
NAmes_plot <- ggplot(NAmes_data, aes(x = lGrLivAreaHundr)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "cornflowerblue") +
  geom_density(aes(color = "Actual")) +
  stat_function(aes(color = "Normal"), 
                fun = dnorm, 
                args = list(mean = mean_Square_Feet, sd = sd_Square_Feet)) +
  ggtitle("Square Feet in North Ames") +
  xlab("Square Feet (in hundreds)") +
  scale_colour_manual("Distribution", values = c("darkred", "darkblue"))

#--------------------------Edwards---------------------------------------

# Calculate mean and standard deviation
mean_Square_Feet <- mean(Edwards_data$lGrLivAreaHundr)
sd_Square_Feet <- sd(Edwards_data$lGrLivAreaHundr)

# Create histogram with normal and density curves (NAMES)
Edwards_plot <- ggplot(Edwards_data, aes(x = lGrLivAreaHundr)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "red") +
  geom_density(aes(color = "Actual")) +
  stat_function(aes(color = "Normal"), 
                fun = dnorm, 
                args = list(mean = mean_Square_Feet, sd = sd_Square_Feet)) +
  ggtitle("Square Feet in Edwards") +
  xlab("Square Feet (in hundreds)") +
  scale_colour_manual("Distribution", values = c("darkred", "darkblue"))

#--------------------------BrkSide--------------------------------
# Calculate mean and standard deviation
mean_Square_Feet <- mean(BrkSide_data$lGrLivAreaHundr)
sd_Square_Feet <- sd(BrkSide_data$lGrLivAreaHundr)

# Create histogram with normal and density curves (BrkSide)
BrkSide_plot <- ggplot(BrkSide_data, aes(x = lGrLivAreaHundr)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "green") +
  geom_density(aes(color = "Actual")) +
  stat_function(aes(color = "Normal"), 
                fun = dnorm, 
                args = list(mean = mean_Square_Feet, sd = sd_Square_Feet)) +
  ggtitle("Square Feet in BrkSide") +
  xlab("Square Feet (in hundreds)") +
  scale_colour_manual("Distribution", values = c("darkred", "darkblue"))

grid.arrange(Combined_plot, NAmes_plot, Edwards_plot, BrkSide_plot, ncol = 2, nrow = 2)
```




```{r unlogged_scatterplot}
library(ggplot2)
ggplot(AmesHousing_Data3, aes(x = GrLivAreaHundr, y = SalePriceHundr, color = Neighborhood)) + 
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE) + 
  labs(title = "Relationship of Price and Living Area by Neighborhood", x = "Total Area", y = "Sale Price", color = "Neighborhood")

```



```{r logged_scatterplot}

library(ggplot2)
ggplot(AmesHousing_Data3, aes(x = lGrLivAreaHundr, y = lSalePriceHundr, color = Neighborhood)) + 
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE) + 
  labs(title = "Relationship of Logged Price and Logged Living Area by Neighborhood", x = "Logged Area (in hundreds of sqrft)", y = "Logged Sale Price (in hundreds)", color = "Neighborhood")
```

```{r conversion}
# Coefficients from the regression model
intercept <- 5.79475
coeff_Edwards <- -0.02507
coeff_NAmes <- 0.12769
coeff_GrLivArea <- 0.53874

# Transforming coefficients back to original scale
original_intercept <- exp(intercept)
original_coeff_Edwards <- exp(coeff_Edwards)
original_coeff_NAmes <- exp(coeff_NAmes)
original_coeff_GrLivArea <- exp(coeff_GrLivArea)

# Interpretation in dollars and square feet
# Assuming GrLivArea is measured in square feet
# Assuming the SalePrice was measured in dollars
# Interpretation of coefficients in original scale
interpretation_intercept <- paste("Intercept (in hundreds of dollars):", round(original_intercept, 2))
interpretation_Edwards <- paste("Effect of Edwards (in hundreds of dollars):", round(original_coeff_Edwards, 2))
interpretation_NAmes <- paste("Effect of NAmes (in hundreds of dollars):", round(original_coeff_NAmes, 2))
interpretation_GrLivArea <- paste("Effect of GrLivArea (in hundreds of square feet):", round(original_coeff_GrLivArea, 2))

# Printing interpretations
print(interpretation_intercept)
print(interpretation_Edwards)
print(interpretation_NAmes)
print(interpretation_GrLivArea)
```






# Initial dataset for Analysis Question 2
```{r housing_regression_large}

AmesHousing_Data$Neighborhood <- as.factor(AmesHousing_Data$Neighborhood)

# Create a new column that holds the'SalePrice' in hundreds
AmesHousing_Data$SalePriceHundr <- floor(AmesHousing_Data$SalePrice /100)

# Create a new column that holds the'SalePrice' in hundreds
AmesHousing_Data$GrLivAreaTen <- floor(AmesHousing_Data$GrLivArea /100)

# Fit a multiple linear regression model
lm_model <- lm(SalePriceHundr ~ GrLivAreaTen + Neighborhood, data = AmesHousing_Data)

# Summarize the model
summary(lm_model)

# Calculate mean and standard deviation
mean_sale_price <- mean(AmesHousing_Data$SalePriceHundr)
sd_sale_price <- sd(AmesHousing_Data$SalePriceHundr)
mean_sqr_ft <- mean(AmesHousing_Data$GrLivAreaTen)
sd_sqr_ft <- sd(AmesHousing_Data$GrLivAreaTen)

# Check Assumptions Visually
plot(lm_model)

# For histograms
ggplot(AmesHousing_Data, aes(x = SalePriceHundr)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "grey") +
  geom_density(aes(color = "Actual")) +
  stat_function(aes(color = "Normal"), 
                fun = dnorm, 
                args = list(mean = mean_sale_price, sd = sd_sale_price)) +
  xlab("Sale Price (in Hundreds)") +
  scale_colour_manual("Distribution", values = c("darkred", "darkblue"))

ggplot(AmesHousing_Data, aes(x = GrLivAreaTen)) +
  geom_histogram(aes(y = ..density..), color = "black", fill = "grey") +
  geom_density(aes(color = "Actual")) +
  stat_function(aes(color = "Normal"), 
                fun = dnorm, 
                args = list(mean = mean_sqr_ft, sd = sd_sqr_ft)) +
  xlab("Area sqr ft (in Hundreds)") +
  scale_colour_manual("Distribution", values = c("darkred", "darkblue"))

# Boxplot for SalePrice 
ggplot(AmesHousing_Data, aes(x = Neighborhood, y = SalePriceHundr)) +
  geom_boxplot() +
  labs(title = "SalePrice Distribution Across Neighborhoods",
       x = "Neighborhood",
       y = "SalePrice (in hundreds)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Boxplot for square feet 
ggplot(AmesHousing_Data, aes(x = Neighborhood, y = GrLivAreaTen)) +
  geom_boxplot() +
  labs(title = "Square Feet Across Neighborhoods",
       x = "Neighborhood",
       y = "Square Feet (in tens)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r examine_all_outliers_large}
# Calculate IQR for SalePrice and GrLivArea
Q1 <- quantile(AmesHousing_Data$SalePrice, 0.25)
Q3 <- quantile(AmesHousing_Data$SalePrice, 0.75)
IQR_SalePrice <- Q3 - Q1

Q1_area <- quantile(AmesHousing_Data$GrLivArea, 0.25)
Q3_area <- quantile(AmesHousing_Data$GrLivArea, 0.75)
IQR_GrLivArea <- Q3_area - Q1_area

# Define thresholds for outliers 
threshold_SalePrice <- 3.5 * IQR_SalePrice
threshold_GrLivArea <- 3.5 * IQR_GrLivArea

# Identify outliers
outliers <- AmesHousing_Data[(AmesHousing_Data$SalePrice < (Q1 - threshold_SalePrice)) |
                              (AmesHousing_Data$SalePrice > (Q3 + threshold_SalePrice)) |
                              (AmesHousing_Data$GrLivArea < (Q1_area - threshold_GrLivArea)) |
                              (AmesHousing_Data$GrLivArea > (Q3_area + threshold_GrLivArea)), ]
outliers

# Assuming 'AmesHousing_Data2' is your dataset
partial_sales <- AmesHousing_Data$SaleCondition == "Partial"
partial_rows <- AmesHousing_Data[partial_sales, ]

partial_rows

# Export outliers to a CSV file
# write.csv(outliers, "D:/University/SMU/Statistical_Foundations_for_Data_Science/Stat_foundations_repository/Statistical_Foundations/Final_Project/outliers.csv", row.names = FALSE)
```



# 131 and 339 in Edwards, have over 10000 sqft including the unfinished basements and 1 and 1.5 acres-most likley an error - no homes over .5 acres built after 2007
# 136 in Brookside, has a price and square feet that are likely missing a zero - Brookside smallest structure has 729 square feet this one, 334
```{r outliers_high_leverage_large}
# Identify the rows of interest'
outlier_indices <- c(339, 136, 131, 169, 190)

# Extracting rows corresponding to the outlier indices
outlier_rows <- AmesHousing_Data2[outlier_indices, ]

outlier_rows


```




